A biometric analysis of infant mortality and temperature, northern Sweden 1895-1950

### 2022-02-15 getdata4.Rmd
fix socst

```{r fromobs}
bobs <- obs[obs$starttyp == 2 & obs$id %in% per$id, ]
indx <- match(per$id, bobs$id)
per$socStatus <- bobs$socStatus[indx]
per$socBranch <- bobs$socBranch[indx]
per$hisclass <- bobs$hisclass[indx]
per$nofrs <- bobs$nofrs[indx]
per <- per[!is.na(per$nofrs), ]
per$subreg <- factor(per$nofrs)
levels(per$subreg) <- c(rep("ume", 2), 
                        rep("bjur", 4), 
                        rep("sten", 3))
##per <- per[(per$subreg != "sten"), ] # NOTE: Now Stensele station is included
per$subreg <- factor(per$subreg)
per$enter <- 0
per$sex <- factor(per$kon, labels = c("boy", "girl"))
per$birthdate <- toTime(per$birthdate)
per$illeg <- as.numeric(per$ab == 2)
per$parity <- per$paritet_g
per$paritet_g <- per$paritet_k <- per$doddat <- per$frsbosdat <- NULL
per$fid <- per$fboort <- per$ab <- per$fyrkfd <- per$fyrkind <- NULL
per$region <- per$kon <- per$fodhfrs <- per$frsbobdat <- per$frsbostyp <- NULL
per$fbonmn <- per$fboind <- per$fyrkrel <- per$fyrketxt <- NULL
per$foddat <- as.Date(as.character(per$foddat), format = "%Y%m%d")
## NEW:
per$socBranch <- as.character(per$socBranch)
per$socBranch[is.na(per$socBranch)] <- "none"
per$socBranch <- factor(per$socBranch)
per$socStatus <- as.character(per$socStatus)
per$socStatus[is.na(per$socStatus)] <- "low"
per$socStatus <- factor(per$socStatus)
per <- per[, c("id", "sex", "birthdate", "mid", "foddat", "parity", "illeg", 
               "subreg", "socBranch", "socStatus", "hisclass", "urban", "enter", "exit", "event")]
saveRDS(per, file = "data/births4.rds")
```

@@

bume$region <- "ume"
bskel$region <- "ske"
bsten$region <- "sten"
infdat <- rbind(bume, bskel, bsten)
infdat$birthmonth <- ceiling((infdat$birthdate - floor(infdat$birthdate)) * 12)
infdat$period <- cut(infdat$birthdate, c(1895, 1914, 1935, 1951), dig.lab = 5)
infdat <- infdat[order(infdat$id, infdat$enter), c("subreg", "id", "sex", "enter", "exit", "event", 
                                                   "birthmonth", "period", "socBranch", "socStatus", "illeg",
                                                   "parity", "hisclass", "urban", tempvars)]
## New addition: (28 Jan 2022)
##infdat <- rc(infdat) # rc from 'skum'
##n <- NROW(infdat)

>>>>

```{r fromobs}
bobs <- obs[obs$starttyp == 2 & obs$id %in% per$id, ]
indx <- match(per$id, bobs$id)
per$socst <- bobs$socst[indx]
per$socst[is.na(per$socst)] <- 9
per$socst1 <- factor(per$socst, labels = c("highBusiness", "highOfficial", "farmer", "lowBusiness",
                                          "lowOfficial", "qualWorker", "farmhand", "unqualWorker", "none"))
per$socst <- factor(per$socst, labels = c("highWhiteC", "highWhiteC", "farmer", "lowWhiteC", "lowWhiteC", 
                                           "qualWorker", "farmer", "worker", "none"))
##per$socStatus <- bobs$socStatus[indx]
##per$socBranch <- bobs$socBranch[indx]
per$hisclass <- bobs$hisclass[indx]
per$nofrs <- bobs$nofrs[indx]
per <- per[!is.na(per$nofrs), ]
per$subreg <- factor(per$nofrs)
levels(per$subreg) <- c(rep("ume", 2), 
                        rep("bjur", 4), 
                        rep("sten", 3))
##per <- per[(per$subreg != "sten"), ] # NOTE: Now Stensele station is included
per$subreg <- factor(per$subreg)
per$enter <- 0
per$sex <- factor(per$kon, labels = c("boy", "girl"))
per$birthdate <- toTime(per$birthdate)
per$illeg <- as.numeric(per$ab == 2)
per$parity <- per$paritet_g
per$paritet_g <- per$paritet_k <- per$doddat <- per$frsbosdat <- NULL
per$fid <- per$fboort <- per$ab <- per$fyrkfd <- per$fyrkind <- NULL
per$region <- per$kon <- per$fodhfrs <- per$frsbobdat <- per$frsbostyp <- NULL
per$fbonmn <- per$fboind <- per$fyrkrel <- per$fyrketxt <- NULL
per$foddat <- as.Date(as.character(per$foddat), format = "%Y%m%d")
## NEW:
##per$socBranch <- as.character(per$socBranch)
##per$socBranch[is.na(per$socBranch)] <- "none"
##per$socBranch <- factor(per$socBranch)
##per$socStatus <- as.character(per$socStatus)
##per$socStatus[is.na(per$socStatus)] <- "low"
##per$socStatus <- factor(per$socStatus)
per <- per[, c("id", "sex", "birthdate", "mid", "foddat", "parity", "illeg", "socst",
               "subreg", "hisclass", "urban", "enter", "exit", "event")]
saveRDS(per, file = "data/births4.rds")
```

@@

bume$region <- "ume"
bskel$region <- "ske"
bsten$region <- "sten"
infdat <- rbind(bume, bskel, bsten)
infdat$birthmonth <- ceiling((infdat$birthdate - floor(infdat$birthdate)) * 12)
infdat$period <- cut(infdat$birthdate, c(1895, 1914, 1935, 1951), dig.lab = 5)
infdat <- infdat[order(infdat$id, infdat$enter), c("subreg", "id", "sex", "enter", "exit", "event", 
                                                   "birthmonth", "period", "socst", "illeg",
                                                   "parity", "hisclass", "urban", tempvars)]
## New addition: (28 Jan 2022)
##infdat <- rc(infdat) # rc from 'skum'
##n <- NROW(infdat)

### 2022-02-15 paper4.Rmd
fix socst

```{r prepare}
source("R/ltx2.R")
infdat$parity <- cut(infdat$parity, c(-2, 1.5, 4.5, 19), labels = c("1", "2-4", "5+"))
##infdat$period <- cut(infdat$birthdate, c(1895, 1914, 1935, 1951), dig.lab = 5)
##levels(infdat$subreg) <- c("ume", "ske")
##infdat$season <- cut(infdat$week, c(-1, 9.5, 22.5, 35.5, 48.5, 52.5), labels = c("winter", "spring", "summer", "fall", "winter"))
##infdat$season <- as.factor(infdat$season)
## New:
infdat$summer <- infdat$week %in% 14:39
infdat$emeantemp <- round(infdat$emeantemp)
levels(infdat$socBranch) <- c("office", "farming", "none", "office", "worker") # NOTE!
```

>>>>

```{r prepare}
source("R/ltx2.R")
infdat$parity <- cut(infdat$parity, c(-2, 1.5, 4.5, 19), labels = c("1", "2-4", "5+"))
##infdat$period <- cut(infdat$birthdate, c(1895, 1914, 1935, 1951), dig.lab = 5)
##levels(infdat$subreg) <- c("ume", "ske")
##infdat$season <- cut(infdat$week, c(-1, 9.5, 22.5, 35.5, 48.5, 52.5), labels = c("winter", "spring", "summer", "fall", "winter"))
##infdat$season <- as.factor(infdat$season)
## New:
infdat$summer <- infdat$week %in% 14:39
infdat$emeantemp <- round(infdat$emeantemp)
```
